---
title: "The rENA Package"
subtitle: "Text Mining Module 4: A Code Along"
format: 
  revealjs:
    # include-in-header: preview.html
    progress: true
    chalkboard: 
      buttons: false
    preview-links: auto
    # embed-resources: true
    logo: img/LASERlogo_large.png
    width: 1920
    height: 1080
    margin: 0.05
    footer: <a href=https://www.fi.ncsu.edu/projects/laser-institute>LASER Institute
    slide-number: c/t
    theme: [default, css/laser.scss]
execute: 
  freeze: true
resources:
  - demo.pdf
editor: 
  markdown: 
    wrap: 72
---

## Welcome to the LAW Code Along for Module 4

- The Text Mining course is designed for those seeking an introductory
  understanding of quantifying the text in documents to better
  understand their properties.

- The following Code Along is a companion to the Module 4 case study's
  **Wrangle** stage.

![Figure 2.2 Steps of Data-Intensive Research
Workflow](img/workflow.png){width="60%"}

[@krumm2018]

::: notes
This Code Along walks through some essential components of using
Epistemic Network Analysis (ENA) to explore collaborative learning
processes in digital STEM education environments.

ENA is a sophisticated technique that visualizes and quantifies
connections among concepts in coded data, offering rich insights into
how learners construct knowledge collaboratively.
:::

## Module Objectives

This Code Along discusses basic features of `rENA`, the ENA package for
R. By the end of this module we will learn how to:

- Import data from
  [RescuShell](https://www.virtualinterns.org/rescushell/), an
  engineering virtual internship platform.
- Specify model parameters for `rENA`: Units, codes, conversations, the
  window, groups, rotations, and the metadata.
- Construct an ENA model.

::: notes
This Code Along walks you through the Wrangle phase of the Case Study.
The setup for this kind of analysis looks like it has a lot of code, but
is no less straightforward than other Code Alongs we've done before.
:::

## Context of the Problem

![Arastoopour Irgens, G., Shaffer, D. W., Swiecki, Z., Ruis, A. R., &
Chesler, N. C. (2015). \[Teaching and assessing engineering design
thinking with virtual internships and epistemic network
analysis\](https://open.clemson.edu/cgi/viewcontent.cgi?article=1004&context=ed_human_dvlpmnt_pub).
International Journal of Engineering
Education.](img/arastoopour.png "Our world is rich with data sources, and technology makes data more accessible than ever before! To help ensure students are future ready to use data for making informed decisions, many countries around the world have increased the emphasis on statistics and data analysis in school curriculum–from elementary/primary grades through college. This course allows you to learn, along with colleagues from other schools, an investigation cycle to teach statistics and to help students explore data to make evidence-based claims. To learn more about engaging learners in making inferences and claims supported by data and how to emphasize inferential reasoning in teaching statistics through posing different types of investigative questions, enroll in our Teaching Statistics through Inferential Reasoning MOOC-Ed."){width="50%"}

::: notes
Paper abstract:

An engineering workforce of sufficient size and quality is essential for
addressing significant global challenges such as climate change, world
hunger, and energy demand. Future generations of engineers will need to
identify challenging issues and design innovative solutions. To prepare
young people to solve big and increasingly global problems, researchers
and educators need to understand how we can best educate young people to
use engineering design thinking. In this paper, we explore virtual
internships, online simulations of 21st-century engineering design
practice, as one method for teaching engineering design thinking. To
assess the engineering design thinking, we use epistemic network
analysis (ENA), a tool for measuring complex thinking as it develops
over time based on discourse analysis. The combination of virtual
internships and ENA provides opportunities for students to engage in
authentic engineering design, potentially receive concurrent feedback on
their engineering design thinking, and develop the identity, values, and
ways of thinking of professional engineers.
:::

## Research Questions:

- What are the patterns of epistemic connections formed by learners as
  they collaboratively engage in engineering problem-solving tasks
  within digital internship environments?

- Is there a difference in learners' epistemic frames based on the
  condition to which they were assigned?

::: notes
These research questions are answered more fully in the case study.
:::

## Load Libraries

::: panel-tabset
## Packages Needed

- `tidyverse`

- `tidytext`

- `rENA`

## 👉 Your Turn ⤵

- Load the `tidyverse`, `tidytext`, and `rENA` packages using
  `library().`

## Answer

```{r echo=TRUE}
library(tidyverse)
library(tidytext)
library(rENA)
```
:::

::: notes
`tidyverse` and `tidytext` are, by now, well-discussed essential
packages for data manipulation and text mining.

The `rENA` package provides a comprehensive set of tools specifically
designed for conducting Epistemic Network Analysis. It enables users to
create visual representations and statistical models of the connections
between coded data elements, particularly useful in educational and
collaborative learning contexts.

Key functions in the `rENA` package facilitate data preparation, network
modeling, and interactive visualization, making it a valuable resource
for researchers aiming to explore complex discourse and interaction
patterns.
:::

## Read in Data

:::::: panel-tabset
## Your Turn

- Read in "rescushell-data.csv" from your Data folder as new object
  `rescushell_data`.

## Answer

```{r}
#| echo: true
rescushell_data <- read_csv("data/rescushell-data.csv")
view(rescushell_data)
```

## Variables

If you look at the object, you'll see a number of variables which we
will be using to set the parameters:

::::: columns
::: {.column width="50%"}
- UserName

- Condition

- CONFIDENCE.Pre

- CONFIDENCE.Post

- CONFIDENCE.Change

- C.Level.Pre

- NewC.Change

- C.Change

- Timestamp

- ActivityNumber
:::

::: {.column width="50%"}
- GroupName

- GameHalf

- GameDay

- text

- Data

- Technical.Constraints

- Performance.Parameters

- Client.and.Consultant.Requests

- Design.Reasoning

- Collaboration
:::
:::::
::::::

::: notes
Our data file consists of discourse from RescuShell, an online learning
simulation where students work as interns at a fictitious company to
solve a realistic engineering design problem in a simulated work
environment. Throughout the internship, students communicate with their
project teams and mentors via online chat, and these chats are recorded
in the text column. A set of qualitative codes were applied to the data
in the “text” column, where a value of 0 indicates the absence of the
code and a value of 1 indicates the presence of the code in a given
line.
:::

## ENA Model Parameters

::: panel-tabset
## Specify Units

- First, we should identify which variables (columns) in the data
  contain identifiers for unique units.

- We will make an object `unitCols` out of the variables `Condition` and
  `UserName.`

```{r}
#| echo: true
unitCols <- c("Condition", "UserName")
```

## Check Object

```{r}
#| echo: true
rescushell_data |>
  select(all_of(unitCols)) |>
  distinct()
```
:::

::: notes
We'll show an example with the first parameter: Units. For this example,
we'll use the `Condition` column and the `UserName` column to define the
units.

In ENA, ***units*** can be individuals, ideas, organizations, or any
other entity whose structure of connections you want to model. To set
the units parameter, specify which column(s) in the data contain the
variables that identify unique units.

To verify that the units are correctly specified, lets subset and
preview the unique values in the units columns. To do that, we'll pipe
`|>` the data frame `rescushell_data`, `select` only the columns whose
names are stored in the character vector `unitCols` using `all_of()`,
which ensures those names are treated literally and throws an error if
any are missing, and then use the `distinct()` function to drop any
duplicate rows across those selected columns.

There should 48 units from two conditions, which means that the ENA
model will produce 48 individual-level networks for each of the units,
and each unit is uniquely associated with either the novice group
(`FirstGame`) or the relative expert group (`SecondGame`).
:::

## ENA Model Parameters, Cont.

::: panel-tabset
## Specify Codes

- The next parameter is identifying the *codes*.

- Codes are researcher-defined concepts whose pattern of association we
  want to model for each *unit***.**

## Your Turn

- Like for *units*, use `select()` to make a new character vector object
  `codeCols` from the following variables: "Data,"
  "Technical.Constraints", "Performance.Parameters",
  "Client.and.Consultant.Requests", "Design.Reasoning", and
  "Collaboration".

- Verify the content of your new object on your raw data with `select()`
  and `all_of()`.

## Answer

```{r}
#| echo: true
#make the codeCols object

codeCols <- c('Data', 'Technical.Constraints', 'Performance.Parameters', 'Client.and.Consultant.Requests', 'Design.Reasoning', 'Collaboration')

#verify the content in your object
rescushell_data |>
  select(all_of(codeCols))
```
:::

::: notes
ENA represents codes as nodes in the networks and co-occurrences of
codes as edges. Most researchers use binary coding in ENA analyses,
where the values in the code columns are either 0 (indicating that the
code is not present in that line) or 1 (indicating that the code is
present in that line). `rescushell_data` contains six code columns, all
of which will be used here.

It's your turn to try making the code parameter, using the same
principle as for the unit parameter. Also try to verify the content in
your object from the raw data using `select()` and `all_of()`. There
isn't a need to use `distinct()` this time.

Your code should return a tibble containing only the six columns you
listed in `codeCols` vector (in exactly that order) and all 3,824 rows
from `rescushell_data`. In other words, you’ll see a preview of the data
frame with just the columns:
:::

## ENA Model Parameters, Cont.

::: panel-tabset
## Conversation

- The *conversation* parameter determines which lines in the data *can*
  be connected. Codes in lines that are not in the same conversation
  cannot be connected.

## Your Turn

- Like for *units*, use `select()` to make a new character vector object
  `conversationCols` from the following variables: "Condition",
  "GroupName", and "ActivityNumber".

- Verify the content of your new object on your raw data with `select()`
  and `all_of()`.

## Answer

```{r}
#| echo: true
#make parameter object
conversationCols <- c("Condition", "GroupName", "ActivityNumber")

#verify
rescushell_data |>
  select(all_of(conversationCols))
```
:::

::: notes
For this case study, we'll choose the `Condition`, `GroupName`, and
`ActivityNumber` columns to define the "conversations." These choices
indicate that connections can only happen between students who were in
the same condition (`FirstGame` or `SecondGame`) and on the same project
team (group), and within the same activity. This definition of
conversation reflects what actually happened in the simulation: in a
given condition, students only interacted with those who were in the
same group, and each activity occurred on a different day.

After verifying, your code should return a tibble containing only the
three columns you listed in `conversationCols` vector (in exactly that
order) and all 3,824 rows from `rescushell_data`.
:::

## ENA Model Parameters, Cont.

::: panel-tabset
## Window

- While the *conversation* parameter specifies which lines **can** be
  related, the window parameter determines which lines within the same
  conversation **are** related.

- The most common window method used in ENA is called a moving stanza
  window.

- We will set the window size to 7 for this case study, using
  `window.size.back`:

```{r}
#| echo: true
window.size.back = 7
```

## Groups & Rotation

- When specifying the units, we chose a column that indicates two
  conditions: `FirstGame` (novice group) and `SecondGame` (relative
  expert group).

- To enable comparison of students in these two conditions, three
  additional parameters need to be specified: `groupVar`, `groups`,
  and `mean` like so:

```{r}
#| echo: true
groupVar <- "Condition" # "Condition" is the column used as our grouping variable 
groups <- c("FirstGame", "SecondGame") # "FirstGame" and "SecondGame" are the two unique values of the "Condition" column
mean = TRUE
```

## Metadata

```{r}
#| echo: true
metaCols = c("CONFIDENCE.Change","CONFIDENCE.Pre","CONFIDENCE.Post","C.Change") # optional
```
:::

::: notes
Once the conversation parameter is specified, a ***window*** method
needs to be specified. Whereas the conversation parameter specifies
which lines *can be* related, the window parameter determines which
lines within the same conversation *are* related. The most common window
method used in ENA is called a moving stanza window, which is what will
be used here.

Briefly, a ***moving stanza window*** is a sliding window of fixed
length that moves through a conversation to detect and accumulate code
co-occurrences in recent temporal context. The lines within a designated
stanza window are considered related to each other. For instance, if the
moving stanza window is 7, then each line in the conversation is linked
to the six preceding lines.

In specifying ***groups*****,** `groupVar`, `groups`, and `mean` are
parameters that indicate the first ENA dimension will maximize the
difference between the two
conditions: `FirstGame` and `SecondGame.` This difference maximization
is achieved through `mean = TRUE`, which specifies that a *means
rotation* will be performed at the dimensional reduction stage.

Finally, the last parameter to be specified is **metadata**. Metadata
columns are not required to construct an ENA model, but they provide
information that can be used to subset units in the resulting model.
:::

## Construct an ENA Model

::: panel-tabset
## Construction

- Now that all the essential parameters have been specified, the ENA
  model can be constructed.

- To build an ENA model, we need two
  functions `ena.accumulate.data` and `ena.make.set`, and we recommend
  that you store the output in an object (in this case, **`set.ena`**).

```{r}
#| echo: true
accum.ena <- 
  ena.accumulate.data(
    text_data = rescushell_data[, 'text'],
    units = rescushell_data[,unitCols],
    conversation = rescushell_data[,conversationCols],
    metadata = rescushell_data[,metaCols], # optional
    codes = rescushell_data[,codeCols],
    window.size.back = 7
)

set.ena = 
  ena.make.set(
    enadata = accum.ena, # the accumulation run above
    rotation.by = ena.rotate.by.mean, # equivalent of mean=TRUE in the ena function
    rotation.params = list(
      accum.ena$meta.data$Condition=="FirstGame", # equivalent of groups in the ena function
      accum.ena$meta.data$Condition=="SecondGame" # equivalent of groups in the ena function
  )
)
```

## Your Turn

- Let's use the `names()` function to first determine what types of
  items and data are stored in the `set.ena` model.

## Answer

```{r}
#| echo: true
names(set.ena)
```
:::

::: notes
This combined visual and numeric approach enables a comprehensive
understanding of discourse patterns within our ENA framework.

Now that we have constructed an ENA Model object, we can begin to
descriptively explore through visual and numerical summaries the data
stored in the `set.ena` model. Take a minute to look at it using
`names()`.
:::

## ❓Questions

- What do you see in the `set.ena` model so far? What does that tell you
  about what we are about to work with in the case study?

Congratulations! You've finished your last Code Along!

# What's Next? {background="#143156"}

::: {style="text-align: left; font-size: 50px"}
- Complete the *Prepare and Wrangle* part of the [case
  study](https://posit.cloud/spaces/493094/join?access_code=qjsumVzA0HZYgrxOSizVbVPmDs7iHYS32yUQaqDe)

- Complete the [Epistemic Network Analysis
  badge](https://laser-institute.github.io//text-mining/module-4/tm-4-badge.html)

- Complete the TM
  [Microcredential](https://laser-institute.github.io//text-mining/module-5/tm-microcredential-badge.html)
:::
