---
title: "What Do Teachers Value About Online PD?"
author: "Shaun Kellogg"
format:
  revealjs: 
    theme: simple
editor: visual
---

## Research Questions

1.  What aspects of online professional learning resources do teachers find most valuable?

2.  How might online learning resources differ in the value they afford teachers?

## Benefits of Online PD

Words teachers frequently used to describe online PD:

```{r}
library(tidyverse)
library(tidytext)
library(wordcloud2)

opd_survey <- read_csv("data/opd_survey.csv")

teacher_counts <- opd_survey |>
  select(Role, Resource, Q21) |> # Select only columns Role, Resource, and Q21
  rename(text = Q21) |> 
  slice(-1, -2) |>
  drop_na() |>
  filter(Role == "Teacher") |>
  unnest_tokens(word, text) |>
  anti_join(stop_words) |>
  count(word, sort = TRUE)

wordcloud2(teacher_counts,
           color = ifelse(teacher_counts[, 2] > 1000, '#CC0000', 'gray'))
```

## Top 15 Words

```{r}
teacher_counts |>
  filter(n > 500) |>
  slice_max(order_by = n, n = 15) |>    
  mutate(word = reorder(word, n)) |>
  ggplot(aes(n, word)) +
  geom_col() + 
  labs(x = "Word Counts", y = NULL) + 
  theme_minimal()
```

## Information Shared

```{r, eval=FALSE}

# This code is used to find quotes, which are copied, pasted, and formatted below

opd_teacher <- opd_survey |>
  select(Role, Resource, Q21) |>
  rename(text = Q21) |>
  slice(-1, -2) |>
  filter(Role == "Teacher")

select(opd_teacher, text) |> 
  filter(grepl('inform*', text)) |>
  sample_n(10) |>
  
```

-   "It could be completed at home in my free time. It had lots of good information that helped me understand the **terminology**."

-   "the information about the way our **common core** was created for our system"

-   "Informative information on the **21st century learner** and teacher."

-   "The most beneficial aspect was the information about **types of data** and where we get it from."

## Videos and Resources Provided

```{r, eval=F}

select(opd_teacher, text) %>% 
  filter(grepl('resource*', text)) %>%
  sample_n(10)

```

-   "**Printable resources** to keep on hand and example videos of how to use formative assessments in the classroom."

-   "I was able to see other **lesson plans** and resources that other schools are doing."

-   "Viewing **teacher videos**, especially Dan Meyer and Finland classroom. I was also introduced to new online resources that will aid my students and me in learning."

## Time and Pace for Learning

```{r, eval=F}

select(opd_teacher, text) %>% 
  filter(grepl('time', text)) %>%
  sample_n(10)

```

-   "The **ability to schedule the time to work** on the module so that it does not interfere with my commitments as an educator."

-   "time to **talk with peers**"

-   "The most valuable aspect was the time factor... I could **walk away and work on it at another time**, so this is a tremendous aspect for the online learner."

-   "Teachers are **able to go back to the information at anytime** to view the information to ensure understanding of the material."

## Top 5 Words by Resource

```{r}

opd_survey |>
  select(Role, Resource, Q21) |> # Select only columns Role, Resource, and Q21
  rename(text = Q21) |> 
  slice(-1, -2) |>
  drop_na() |>
  filter(Role == "Teacher") |>
  unnest_tokens(word, text) |>
  anti_join(stop_words) |>
  count(Resource, word, sort = TRUE) |>
  group_by(Resource) |>
  mutate(percent = n / sum(n)) |>
  filter(Resource != "Calendar") |> 
  group_by(Resource) |> 
  slice_max(order_by = percent, n = 5, with_ties = FALSE) |> 
  ungroup() |>  
  mutate(word = reorder_within(word, percent, Resource)) |>  
  ggplot(aes(percent, word, fill = Resource)) + 
  geom_col(show.legend = FALSE) +
  facet_wrap(~Resource, ncol = 3, scales = "free_y") +  
  scale_y_reordered()
```

## Conclusions

1.  Teachers found the **information** shared beneficial, such as new terminology, the common core standards, and types of data.
2.  Teachers appreciated the **resources and videos** provided including lesson plans and printable materials.
3.  Use of words like time, pace, and format suggests teachers valued the **convenience** of online professional development.

## Discussion

-   **Limitations**: While able to provide quick insight into a large number of responses, analysis raised as many questions as it answered.
-   **Implications**: Validation of design decisions such use of videos and inclusion of practical resources.
-   **Next Steps for Analysis**: Dig into recommendations for improvement.
